---
title: "Challenge 7: Incorporating Multiple Inputs"
subtitle: "Functions + Fish"
author: "Skylar Richards"
format: html
editor: source
embed-resources: true
message: false
warning: false
code-fold: true
code-tools: true
---

```{r}
#| label: setup
#| message: false
library(tidyverse)
fish <- read.csv("data/BlackfootFish.csv")

```

## Condition Index

A frequently used measurement for fish health is a condition index [(Wikipedia article)](https://en.wikipedia.org/wiki/Condition_index_in_fish). The following
simple equation can be used to calculate the approximate condition index of a
fish:

$$\text{condition index} = \frac{weight}{length^3} \times 100$$

**1. There are specific units required for the calculation of a condition index -- length must be in centimeters and weight must be in grams. The weight data for the Blackfoot River fish were collected in grams; the length data were collected in millimeters. Transform the length data to the correct units.**

```{r}
#| label: transform-data-to-have-correct-units
fish_challenge <- fish|>
  mutate(length = length/10)
```

**2. Collecting data of this sort can be very messy! Write a function that will replace unlikely `length` and `weight` measurements with `NA`. Your function should accept at least three inputs:**

+ **a vector of measurements,**
+ **the minimum reasonable value,**
+ **the maximum reasonable value.**

**If a value falls outside these bounds, you should replace it with an `NA`.**

::: callout-tip
If you are struggling with the structure of your function, I would suggest reading the [Mutating Function](https://r4ds.hadley.nz/functions.html#mutate-functions) from R4DS.
:::

```{r}
#| label: function-to-change-impossible-measurements-to-NA
outlier_deletion <- function(variable_interest, min_reasonable, max_reasonable) {
  variable_interest <- ifelse(variable_interest < min_reasonable | variable_interest > max_reasonable, NA, variable_interest)
}
```

**3. After consulting the [Montana Record Table](https://fwp.mt.gov/fish/anglingData/records) for the four species of trout included in these data, I have conjectured that it is unlikely to have measurements for fish below 5 cm and above 80 cm in length or below 10 g and above 4,000 g in weight. Use your function to modify the `length` and `weight` columns of the `fish` dataset based on my cutoffs.**

```{r}
#| label: modify-impossible-values-of-length-weight
fish_no_outlier <- fish_challenge|>
  mutate(
    length = outlier_deletion(length, 5, 80),
    weight = outlier_deletion(length, 10, 4000)
  )
```

**3. Write a function to calculate the condition index of a fish, given inputs of weight and length.**

::: callout-warning
Consider whether you will write a function that works with vectors (accepting
vectors as inputs) or a function that works with data frames (accepting a 
dataset and variable names as inputs)!
:::

```{r}
#| label: condition-function
#condition_indexer <- function(dataframe, weight, length){
 # dataframe|>
 #   mutate(condition_index = (({{weight}}/({{length}}^3))*100))
#}
condition_indexer <- function(dataframe, weight, length) {
  dataframe |>
    mutate(condition_index = (({{ weight }} / ({{ length }}^3)) * 100))
}

conditioned_fish <- fish_no_outlier|>
  condition_indexer(weight, length)
```

**4. Make a thoughtful visualization of how fish conditions have varied over the duration of this study.**

::: callout-warning
### To be thoughtful...

... you need to use at least a third variable!
:::

```{r}
#| label: condition-indices-over-time
#| 
ggplot(conditioned_fish,
       aes(x = year, y = condition_index, color = species))+
  geom_point(alpha = 0.1)+
  geom_smooth(method = "loess", se = FALSE, aes(color = species))+
  facet_wrap(~ section)+
  labs(
    title = "Condition Index by year, section, and species",
    x = "Year",
    y = ""
  )
```
